---
title: "Data Preparation"
author: "Satoshi, Nora, Adam, Jiyoon"
date: "December 13, 2018"
output: html_document
---

```{r, message = FALSE, warning = FALSE}
library(dplyr)
library(leaflet)
library(geosphere)
library(data.table)
library(ggplot2)
library(lubridate)
library(caret)
library(readxl)
```

# Part 1: Data preparation

We used all of the bluebike data in 2017 

## Read the bluebike data in 2017
```{r}
# Importing the all monthly data in 2017
dat_201701<-read.csv("201701-hubway-tripdata.csv",stringsAsFactors = FALSE)
dat_201702<-read.csv("201702-hubway-tripdata.csv",stringsAsFactors = FALSE)
dat_201703<-read.csv("201703-hubway-tripdata.csv",stringsAsFactors = FALSE)
dat_201704<-read.csv("201704-hubway-tripdata.csv",stringsAsFactors = FALSE)
dat_201705<-read.csv("201705-hubway-tripdata.csv",stringsAsFactors = FALSE)
dat_201706<-read.csv("201706-hubway-tripdata.csv",stringsAsFactors = FALSE)
dat_201707<-read.csv("201707-hubway-tripdata.csv",stringsAsFactors = FALSE)
dat_201708<-read.csv("201708-hubway-tripdata.csv",stringsAsFactors = FALSE)
dat_201709<-read.csv("201709-hubway-tripdata.csv",stringsAsFactors = FALSE)
dat_201710<-read.csv("201710-hubway-tripdata.csv",stringsAsFactors = FALSE)
dat_201711<-read.csv("201711-hubway-tripdata.csv",stringsAsFactors = FALSE)
dat_201712<-read.csv("201712-hubway-tripdata.csv",stringsAsFactors = FALSE)

# Combine them
dat_2017 <- rbind(dat_201701, dat_201702, dat_201703, dat_201704, dat_201705,
                  dat_201706, dat_201707, dat_201708, dat_201709, dat_201710,
                  dat_201711, dat_201712)
```

## Add variables

### 1. age, age_cat, duration_min, year, month, month_abb, day, hour, wday, weekend
```{r}
bbike <- dat_2017 %>%
  mutate(birth.year = as.numeric(birth.year)) %>%
  mutate(age = 2017 - birth.year) %>%
  mutate(age_cat = case_when(
    .$age >= 10 & .$age < 20 ~ 1,
    .$age >= 20 & .$age < 30 ~ 2,
    .$age >= 30 & .$age < 40 ~ 3,
    .$age >= 40 & .$age < 50 ~ 4,
    .$age >= 50 & .$age < 60 ~ 5,
    .$age >= 60 & .$age < 70 ~ 6,
    .$age >= 70 & .$age < 80 ~ 7,
    .$age >= 80 ~ 8)) %>%
  mutate(duration_min = tripduration / 60) %>%
  mutate(year = year(starttime), 
         month = month(starttime),
         month_abb = month(starttime, label = TRUE, abbr = TRUE), 
         day = day(starttime),
         hour = hour(starttime), 
         wday = wday(starttime, label = TRUE, abbr = TRUE))
```

### 2. dist_km
```{r}
setDT(bbike)[ , dist_km := distGeo(matrix(c(start.station.longitude, start.station.latitude), ncol = 2),matrix(c(end.station.longitude, end.station.latitude), ncol = 2))/1000]
bbike <- as.data.frame(bbike)
```

### 3. overtime (if duration_min >  45, 1, 0)
```{r}
bbike <- bbike %>%
  mutate(overtime = ifelse(duration_min > 45, "1", "0"))
```

### 4. user_start, user_end: number of users at the start/end station
```{r}
bbike <- bbike %>%
  group_by(start.station.id) %>%
  mutate(user_start = n())

bbike <- bbike %>%
  group_by(end.station.id) %>%
  mutate(user_end = n())
```

## Combine weather information: temp_max, temp_min, rain, snownice(snow or ice)
```{r}
weather <- read_excel("boston_weather.xls")
```

```{r}
bbike <- bbike %>%
  group_by(year, month, day) %>%
  left_join(., weather, by = c("year", "month", "day")) 
```

## Machine Learning - using logistic regression to predict overtime users characteristics 

### Rationale: With an annual membership, users can ride unlimited bike ride for up to 45 mins. Additional $2.5 will be charged per 30 mins for members. In this analysis, we are interested in learning about the characteristics of overtime riders so that Bluebike can gain extra revenue. 


## Only keep memebers in the dataset 

```{r}
bbikemember <- bbike %>%
  filter(usertype == "Subscriber")

```


### logistic regression
```{r}
set.seed(1)

library(caret)

Train <- createDataPartition(bbikemember$overtime, p=0.8, list=FALSE)
training <- bbikemember[ Train, ]
testing <- bbikemember[ -Train, ]

bbikemember $overtime = as.factor(as.numeric(as.character(bbikemember $overtime)))

#overall accuracy 
p = 0.009  #9892/1102560
y_hat <- sample(c("0","1"), length(testing), replace = TRUE, prob=c (p, 1-p)) %>% 
  factor(levels = levels(bbikemember$overtime))
mean(y_hat == testing$overtime)

#logistic regression
glm.fit <- bbikemember %>% 
  glm( overtime ~  gender + age + month + day + hour + wday + rain + snownice, data=., family = "binomial")

summary(glm.fit)

#wday.L = Sunday wday.Q = Mondaywday.C = Tuesdaywday^4 = Wednesday wday^5 = Thursday wday^6 = Friday Intercept = Saturday 

#prediction
p_hat <- predict(glm.fit, newdata=testing,type="response")
y_hat <- ifelse(p_hat > 0.5, "1", "0") %>% factor()

#confusion matrix 
confusionMatrix(y_hat, testing$overtime)
table(predicted = y_hat, actual = testing$overtime)

```


#### ROC  
```{r}

library(purrr)
library(caret)
library(ggplot2)

probs <- seq(0, 1, length.out = 10)
guessing <- map_df(probs, function(p){
  y_hat <- 
    sample(c("0", "1"), length(testing), replace = TRUE, prob=c(p, 1-p)) %>%
    factor(levels = levels(bbikemember$overtime))
  list(method = "Guessing",
       FPR = 1 - specificity(y_hat, bbikemember$overtime),
       TPR = sensitivity(y_hat, bbikemember$overtime))
})
guessing %>% qplot(FPR, TPR, data =., xlab = "1 - Specificity", ylab = "Sensitivity")

```

